Block bootstrap consistency\\under weak assumptions

### 2017-11-05 Paper.tex
Make minor text changes requested for resubmission

\maketitle

\begin{abstract}\noindent
  This paper weakens the size and moment conditions needed for typical
  block bootstrap methods (i.e.\ the Moving Blocks, Circular Blocks,
  and Stationary Bootstraps) to be valid for the sample mean of
  Near-Epoch-Dependent (\ned) functions of mixing processes; they are
  consistent under the weakest conditions that ensure the original
  \ned\ process obeys a Central Limit Theorem \citep[those
    of][\textit{Econometric Theory}]{Jon:97}.  In doing so, this paper
  extends De Jong's method of proof, a blocking argument, to hold with
  random and unequal block lengths.  This paper also proves that
  bootstrapped partial sums satisfy a Functional \clt\ under the same
  conditions.

@@

\newpage
\noindent Block bootstraps, e.g.\ the Moving Blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary
  Bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$, although, as we will show, their
stronger assumption is necessary for the bootstrap to be consistent
under some forms of heterogeneity.

The proof exploits the conditional independence of the blocks in each
bootstrap.  Each bootstrap proceeds by drawing blocks of $M$
consecutive observations from the original time series, and then
pasting these blocks together to create the new bootstrap time series.
The Moving Blocks bootstrap does exactly that; the Circular Block
bootstrap ``wraps'' the observations, so that $(X_{n-1}, X_n, X_1,
X_2)$, for example, is a possible block of length four (letting $X_t$
denote the original time series).  The Stationary Bootstrap wraps the
observations and also draws $M$ at random for each block;
\citet{PoR:94} suggest drawing $M$ from the geometric distribution.
As the name suggests, the series produced by the Stationary Bootstrap
are strictly stationary, while those produced by the other methods are
not.  Although the Stationary Bootstrap was believed to be much less
efficient than other block bootstrap methods due to results of
\citet{Lah:99}, \citet{Nor:09} has shown that it is only slightly less
efficient than the other block-bootstrap methods discussed in this
paper, and has efficiency identical to that of the non-overlapping
block bootstrap.  Consequently, there has been renewed interest in the
Stationary Bootstrap since stationarity of the bootstrap samples can
be a useful property in theoretical research.  \cite{KrP:11} provides
a recent review of the bootstrap for time-series
processes%
\footnote{Also see the discussion papers by \citet{Dah:11},
\citet{GoP:11}, \citet{Hor:11}, \citet{JeM:11}, and
\citet{KrP:11b}.} %
and \citet{GoP:11} further discuss recent developments in
block-bootstraps.

Theorem~\ref{T1} presents the main result, asymptotic normality of
the distribution of bootstrapped sums. This paper adopts the
standard notation that $\E^{*}$, $\var^{*}$, etc.\ are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation
for precision.  Also note that all results are presented for the
scalar case but generalize immediately to random vectors.  All of the
proofs are presented in the appendix; only proofs for the Stationary
Bootstrap are presented, since proofs for the other methods are
similar and easier to construct.  All limits are taken as $n \to \infty$
unless otherwise noted and $\lVert \cdot \rVert_r$ denotes the
$L_r$-norm.

\begin{thm}\label{T1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n p_n^2)$ (for
    the Stationary Bootstrap) or
    $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n/M_n^2)$ (for
    the Moving or Circular Block Bootstraps),
    where $\E X_{nt} = \mu_{nt}$, $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$, and $p_n$ and $M_n$ are further defined in item 4.
    Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\sigma^{*2} \to^p \sigma^2$, $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

As mentioned earlier, these are the same size and mixing conditions
used by \citet{Jon:97}.
Note that De Jong does allow a little bit more flexibility in the
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows for a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires this condition to hold for every possible partition of
blocks.  This additional restriction is required because the
Stationary Bootstrap will select the blocks randomly and is similar to
\citepos{JoD:00b} requirement for the \fclt.

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller
than they require.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$:
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}.
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have introduced $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and
the same relationship must hold for the expected block length for the
Stationary Bootstrap. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, \pi (1 - \pi) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that $\pi$ could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent and can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L4} and~\ref{L6} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

Theorem~\ref{T1} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{T2} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{T1}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \E^* \bar X_n^*).
\end{align*}
Also let $W$ denote standard Brownian Motion and $\sigma W$ denote
Brownian Motion scaled by the constant $\sigma$.

\begin{thm}\label{T2}
Suppose that the conditions of Theorem~\ref{T1} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive  $\delta$.
If, in addition, $\sup_{t=1,\dots,n}|\mu_{nt} -
\bar \mu_n| = o(1/\sqrt{n})$ and
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{T2} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian Motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian Motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

If the original process does not satisfy~\eqref{eq:4}, it would be
necessary to normalize $W_n$ with a uniformly
consistent estimator of the true covariance process of the series to
make use of these results. This would be the case if the variance
permanently changes partway through the series, for example.  Other
methods, such as the Local Block Bootstrap \citep{PaP:02,DPP:03}, may
be able to capture this additional heterogeneity with the bootstrap
alone, but we do not pursue that possibility further.

@@

To prove~\eqref{eq:19}, it suffices (via the LIE and Lemma~\ref{L2})
to show that
\begin{equation*}
  \sum_{j=1}^{J_n} \Em \big ( Z_{nj}^{*2} \ind \{ Z_{nj}^{*2} > \epsilon \} \big ) \to^p 0.
\end{equation*}
But note that
\begin{equation*}
  \sum_{j=1}^{J_n} \Em \big ( Z_{nj}^{*2} \ind \{ Z_{nj}^{*2} > \epsilon \} \big )
  \leq \max_{j=1,\dots,J_n} \Em \big ( \tfrac{ n }{ M_{nj} } Z_{nj}^{*2}
  \ind \{ \tfrac{ n }{ M_{nj} } Z_{nj}^{*2} >
  \min_{k=1,\dots,J_n} \tfrac{ n }{ M_{nk} } \epsilon \} \big )
\end{equation*}
almost surely. Lemmas~\ref{L1} and~\ref{L6} imply that the RHS converges to
zero in probability as $n \to \infty$,
proving~\eqref{eq:19}. The Dominated Convergence Theorem and~(\ref{eq:18})
then ensure that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
(Also see
Lemma~\ref{L2}.)

>>>>

\maketitle
\nocite{Jon:97}

\begin{abstract}\noindent
  This paper weakens the size and moment conditions needed for typical
  block bootstrap methods (i.e.\ the moving blocks, circular blocks,
  and stationary bootstraps) to be valid for the sample mean of
  Near-Epoch-Dependent (\ned) functions of mixing processes; they are
  consistent under the weakest conditions that ensure the original
  \ned\ process obeys a central limit theorem (CLT),
    established by \citet[\textit{Econometric Theory}]{Jon:97}.  In doing so, this paper
  extends De Jong's method of proof, a blocking argument, to hold with
  random and unequal block lengths.  This paper also proves that
  bootstrapped partial sums satisfy a functional \clt\ (\fclt) under the same
  conditions.

@@

\newpage
\noindent Block bootstraps, e.g.\ the moving blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), circular block \citep{PoR:92}, and stationary
bootstraps \citep{PoR:94}, have become popular in economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the moving blocks
  bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the stationary
  bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$, although, as we will show, their
stronger assumption is necessary for the bootstrap to be consistent
under some forms of heterogeneity.

The proof exploits the conditional independence of the blocks in each
bootstrap.  Each bootstrap proceeds by drawing blocks of $M$
consecutive observations from the original time series, and then
pasting these blocks together to create the new bootstrap time series.
The moving blocks bootstrap does exactly that; the circular block
bootstrap ``wraps'' the observations, so that $(X_{n-1}, X_n, X_1,
X_2)$, for example, is a possible block of length four (letting $X_t$
denote the original time series).  The stationary bootstrap wraps the
observations and also draws $M$ at random for each block;
\citet{PoR:94} suggest drawing $M$ from the geometric distribution.
As the name suggests, the series produced by the stationary bootstrap
are strictly stationary, while those produced by the other methods are
not.  Although the stationary bootstrap was believed to be much less
efficient than other block bootstrap methods due to results of
\citet{Lah:99}, \citet{Nor:09} has shown that it is only slightly less
efficient than the other block-bootstrap methods discussed in this
paper, and has efficiency identical to that of the non-overlapping
block bootstrap.  Consequently, there has been renewed interest in the
stationary bootstrap since stationarity of the bootstrap samples can
be a useful property in theoretical research.  \cite{KrP:11} provides
a recent review of the bootstrap for time-series
processes%
\footnote{Also see the discussion papers by \citet{Dah:11},
\citet{GoP:11}, \citet{Hor:11}, \citet{JeM:11}, and
\citet{KrP:11b}.} %
and \citet{GoP:11} further discuss recent developments in
block-bootstraps.

Theorem~\ref{T1} presents the main result, asymptotic normality of
the distribution of bootstrapped sums. This paper adopts the
standard notation that $\E^{*}$, $\var^{*}$, etc.\ are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation
for precision.  Also note that all results are presented for the
scalar case but generalize immediately to random vectors.  All of the
proofs are presented after the main text of the paper; only proofs for
the stationary bootstrap are given, since proofs for the other methods
are similar and easier to construct.  All limits are taken as
$n \to \infty$ unless otherwise noted and $\lVert \cdot \rVert_r$
denotes the $L_r$-norm.

\begin{thm}\label{T1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $\{V_{nt}\}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded
    and $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n p_n^2)$ (for
    the stationary bootstrap) or
    $\sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2 = o(n/M_n^2)$ (for
    the moving or circular block bootstraps),
    where $\E X_{nt} = \mu_{nt}$, $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$, and $p_n$ and $M_n$ are further defined in item 4.
    Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$, with $\bar X_n = (1/n) \sum_{t=1}^n X_{nt}$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the stationary bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the moving or circular block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$. The last block, $M_{n,J_n}$, is defined as
    $K_{n,J_n} - K_{n,J_n-1}$, so $K_{n,J_n} = n$ a.s.
  \end{enumerate}
  Then $\sigma^{*2} \to^p \sigma^2$, $\hat\sigma^{*2} \to^p \sigma^2$,
  \begin{equation}\label{eq:36}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \E^* \bar X_n^*) \leq x \big]
    - \pr \big[\sqrt{n}(\bar X_{n} - \E \bar X_n) \leq x \big] \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:2}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} -  \E^* \bar X_n^*) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \allcaps{cdf} of the Standard Normal
  distribution and
    \begin{equation}
      \label{eq:3}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

As mentioned earlier, these are the same size and mixing conditions
used by \citet{Jon:97}.
Note that De Jong does allow a little more flexibility in the
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows for a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires this condition to hold for every possible partition of
blocks.  This additional restriction is required because the
stationary bootstrap will select the blocks randomly and is similar to
\citepos{JoD:00b} requirement for the \fclt.

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller
than they require.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$:
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}.
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have introduced $D_n$ to minimize notation in the subsequent
discussion.  For the moving and circular block bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and
the same relationship must hold for the expected block length for the
stationary bootstrap. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, \pi (1 - \pi) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that $\pi$ could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the moving blocks bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent and can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L4} and~\ref{L6} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the circular and stationary
bootstraps.

@@

Theorem~\ref{T1} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{T2} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{T1}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \E^* \bar X_n^*).
\end{align*}
Also let $W$ denote standard Brownian motion and $\sigma W$ denote
Brownian motion scaled by the constant $\sigma$.

\begin{thm}\label{T2}
Suppose that the conditions of Theorem~\ref{T1} hold
and let $\dist$ be any distance function that metricizes weak
convergence. Then
\begin{equation}\label{eq:7}
\pr^{*}[\dist(W_n^{*}, \sigma W) > \delta] \to^p 0
\end{equation}
for any positive  $\delta$.
If, in addition, $\sup_{t=1,\dots,n}|\mu_{nt} -
\bar \mu_n| = o(1/\sqrt{n})$ and
\begin{equation}\label{eq:4}
n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \cov(X_{ns}, X_{nt}) \to \sigma^2 \gamma
\end{equation}
for all $\gamma \in [0,1]$, then
\begin{equation}\label{eq:8}
\pr[\dist(W_n, \sigma W) > \delta] \to 0
\end{equation}
for any positive $\delta$.
\end{thm}
If both~\eqref{eq:7} and~\eqref{eq:8} hold then the bootstrap can be
used to approximate the distribution of partial sums.
Note that Theorem~\ref{T2} imposes
stronger assumptions for the original process
than for the bootstrapped process. Without~\eqref{eq:4}, the
partial sum of the original observations converges to a transformed
Brownian motion with a different covariance process. The bootstrapped
partial sum, on the other hand, always converges to standard (but
potentially scaled) Brownian motion because the resampling strategies
ensure that the bootstrapped process is globally covariance stationary.

If the original process does not satisfy~\eqref{eq:4}, it would be
necessary to normalize $W_n$ with a uniformly
consistent estimator of the true covariance process of the series to
make use of these results. This would be the case if the variance
permanently changes partway through the series, for example.  Other
methods, such as the local block bootstrap \citep{PaP:02,DPP:03}, may
be able to capture this additional heterogeneity with the bootstrap
alone, but we do not pursue that possibility further.

@@

To prove~\eqref{eq:19}, it suffices (via the law of iterated expectations and Lemma~\ref{L2})
to show that
\begin{equation*}
  \sum_{j=1}^{J_n} \Em \big ( Z_{nj}^{*2} \ind \{ Z_{nj}^{*2} > \epsilon \} \big ) \to^p 0.
\end{equation*}
But note that
\begin{equation*}
  \sum_{j=1}^{J_n} \Em \big ( Z_{nj}^{*2} \ind \{ Z_{nj}^{*2} > \epsilon \} \big )
  \leq \max_{j=1,\dots,J_n} \Em \big ( \tfrac{ n }{ M_{nj} } Z_{nj}^{*2}
  \ind \{ \tfrac{ n }{ M_{nj} } Z_{nj}^{*2} >
  \min_{k=1,\dots,J_n} \tfrac{ n }{ M_{nk} } \epsilon \} \big )
\end{equation*}
almost surely. Lemmas~\ref{L1} and~\ref{L6} imply that the RHS converges to
zero in probability as $n \to \infty$,
proving~\eqref{eq:19}. The Dominated Convergence Theorem and~(\ref{eq:18})
then ensure that
\begin{equation}\label{eq:22}
  \pr^{*} \Big[\sum_{j=1}^{J_n}
  Z_{nj}^{*} / \sigma_n^* \leq x \Big] - \Phi(x) \to^p 0.
\end{equation}
(Also see
Lemma~\ref{L2}.)
